\secrel{Abstract Superclasses}

We did not discuss the classes Collection, KeyedCollection,
SequenceableCollection, or ArrayedCollection in the last section, even
though they were listed as forms of "collection" in the beginning of this
chapter. These classes, along with such classes as Boolean, Magnitude,
or Number, are what are known as abstract superclasses. Instances of
abstract superclasses are seldom useful by themselves, but the classes are
important in providing methods that can be inherited by subclasses. For
example, while it is legal in Smalltalk to say:
\begin{lstlisting}
x Collection new
\end{lstlisting}
and the resulting variable x will indeed be an instance of class Collection,
the object is not particularly useful. It has no insertion or deletion protocol,
for example. An instance of class Set, however, is very useful, and the
messages defined in class Collection are important in providing functionality to objects of this class and to other subclasses of class Collection.

The selection and design of abstract superclasses is one of the more
important arts in Smalltalk programming. For example, if one were designing a system to manipulate banking accounts, a class Account might be
a useful abstract superclass for classes CheckingAccount and
SavingsAccount. The actions specific to the individual types of accounts
would be in the subclasses, whereas any common behavior, such as the
actions necessary for opening and closing an account or querying the
balance, might be implemented in the superclass.
