\secrel{Primitives, Cascades, and Coercions}\secdown

This chapter introduces the syntax for cascaded expressions and describes the
notion of primitive expressions. It illustrates the use of primitives by showing
how primitives are used to produce the correct results for mixed mode arithmetic
operations.

\input{language/cascades}
\input{language/prims}
\input{language/numbers}

\secly{Exercises}

\begin{enumerate}

    \item One advantage cited for cascaded expressions was that several
Smalltalk statements could be combined together into one expression. For example
the initialized Bag discussed in the text could be used as an argument to
another object as follows:
\begin{lstlisting}
anObject foo: (Bag new; add: 0; add 1)
\end{lstlisting}
How else might this be done in a single expression without using cascading? You
can use temporary variables, if you wish.

\item What will the result of typing the following expression be? Explain why.
\begin{lstlisting}
2 + (3 print) ; + 4
\end{lstlisting}

\item Examine the class descriptions for the classes Object, Magnitude, Number,
Integer, and Float. Explain in detail how radians respond to the messages < =
and > = .

\item Rewrite the methods for = and < in class Radian so that radians can be
compared only to other radians.

\item Does the class Integer need to provide a method for coerce:? Why or why
not? How about the class Number?

\item An alternative to having instances of class Radian maintain their value in
an instance variable would be to make the class a subclass of class Float.
Discuss the advantages and disadvantages of this approach.

\item Recall that in the hierarchy of number classes all other classes have a
higher ranking then any of Integer or Float. Assume that the class Number
contains the following method
\begin{lstlisting}
    i 
        Complex new; imagpart: self
\end{lstlisting}
That is, in response to the message i, a number will create a new instance of
the type Complex and initialize it with the current object.

Using this method, define the class Complex used to manipulate complex numbers.
Your class should implement methods for the following messages:

\noindent
\begin{tabular}{l p{8cm}}
    new & Set both the imaginary and real portions of the complex number to
    zero. \\
    realpart: & Define the real part of the current number to be the argument.
    \\
    imagpart: & Define the imaginary part of the current number to be the
    argument. \\
    coerce: & Coerce a non-complex, returning an equivalent complex number with
    zero imaginary part. \\
    + & Complex addition. \\
    * & Complex multiplication. \\
    printString & produce a printable representation of the number. \\
\end{tabular}

\noindent
Test your class description by typing several example expressions involving
complex numbers.

\item A useful control structure in many programming languages is a multiway
switch statement, which permits the selection of one out ofseveral possibilities
based on the value of some selection key. Using cascades and blocks we can
implement a multiway switch in Smalltalk. The class Switch will use the message
new: both as a creation message and to assign the switch selection key. The
message case:do: will compare the first argument to the selection key, and, if
equal, the second argument (a block) will be evaluated. The message test:do:
uses blocks for both arguments. The first argument, a one parameter block, is
evaluated using the selection key; if it returns true, the second argument is
evaluated. A flag is maintained by each instance of Switch indicating whether
any condition has been satisfied. The message default: executes the argument, a
block, if no previous condition has been met.

For example, the following statement uses a variable suit representing the suit
of a card from a deck of cards. It places into color a string representing the
color ofthe card.
\begin{lstlisting}
    Switch new: suit;
    case: #spade do: [ color~lblack'] ;
    case: #c1ub do: [ color~lblack' ] ;
    test: [:x I(x = #diamond) or: [x = #heart] ] do: [ color~'red'] ;
    default: [ self error: 'unknown suit I , suit]
\end{lstlisting}
Provide a class description for Switch.

\end{enumerate}

\secup
