\secrel{Basic Classes}\secdown

Основные классы, включенные в стандартную библиотеку Little Smalltalk, описаны в этой главе.

\bigskip

The classes included in the Little Smalltalk system can be roughly divided
into four overlapping groups. These groups are Basic Objects (Object,
UndefinedObject, Symbol, Boolean, True, False, Magnitude, Number,
Char, Integer, Float, Radian, Point), Collections (Collection, Bag, Set,
SequenceableCollection, KeyedCollection, Dictionary, Interval, List,
ArrayedCollection, Array, String, File, Random, ByteArray), Control
Structures (Boolean, True, False, Interval, Block), and System Management (Object, Class, Smalltalk, Process). The following sections will
briefly describe each of these categories. Appendix 3 provides detailed de;.
scriptions for each of the standard classes.

\input{language/basobj}
\input{language/collections}
\input{language/control}
\input{language/classman}
\input{language/abstract}

\secly{EXERCISES}

\begin{enumerate}

\item Suppose you created a new instance of the class UndefinedObject, as
follows:
\begin{lstlisting}
    I~UndefinedObject new
\end{lstlisting}
How does i respond to print? To isNil or notNil? To the object comparison message = = with nil ? Is i nil? List facts to support your
answer.

\item Note that the messages arcSin and arcCos produce an object of type
Radian and not oftype Float. Furthermore, only objects oftype Radian
respond to sin and cos. An alternative would have been to eliminate
the class Radian and to permit all objects of class Float to respond
to the messages sin and cos. Discuss the advantages and disadvantages
of these two different arrangements.

\item Suppose you have a Bag containing numbers. How would you go about
producing an instance of the class List containing the numbers listed
in sorted order?

\item What is the class of Class? what is the superclasss of Class?

\item Many times, two ormore different sequences ofmessages to collections
will have the same effect. In each of the following, describe a sequence
of messages that will have the same effect ,\S the indicated message.
\begin{enumerate}
    \item implement reject: in terms of select:
    \item implement size in terms of inject:into:
    \item implement includes: in terms of inject:into:.
\end{enumerate}

\end{enumerate}

\secup
