\clearpage
\secly{Предисловие}

\input{language/history}
\input{language/about}

\subsecly{Благодарности}

I am, of course, most grateful to the students in the graduate seminar at the
University of Arizona where the Little Smalltalk system was developed. The many
heated discussions and insightful ideas generated were most enjoyable and
stimulating. Participants in that seminar were Mike Benhase, Nick Buchholz, Dave
Burns, John Cabral, Clayton Curtis, Roger Hayes, Tom Hicks, Rob McConeghy,
Kelvin Nilsen, May Lee Noah, Sean O'Malley, and Dennis Vadner. This text grew
out of notes developed for that course, and includes many ideas contributed by
the participants. In particular I wish to thank Dave Burns for the original
versions of the simulation described in Chapter 7 and Mike Benhase and Dennis
Vadner for their work on processes and the dining philosophers solution
presented in Chapter 10.

Discussions with many people have yielded insights or examples that eventually
found their way into this book. I wish to thank, in particular, Jane Cameron,
Chris Fraser, Ralph Griswold, Paul Klint, Gary Levin, and Dave Robson.

Irv Elshoff provided valuable assistance by trying to learn Smalltalk from an
early manuscript and by making many useful and detailed comments on the text.

J. A. Davis from Iowa State University, Paul Klint from the CWI, David Robson
from Xerox Palo Alto Research Center, and. Frances Van Scoy from West Virginia
University provided careful and detailed comments on earlier drafts of the book.

Charlie Allen at Purdue, Jan Gray at Waterloo and Charles Hayden at AT\&T were
early non-Arizona users of Little Smalltalk and were extremely helpful in
finding bugs in the earlier distributions.

I wish to thank Ralph Griswold, Dave Hanson, and Chris Fraser, all chairmen of
the computer science department at the University of Arizona at various times in
the last five years, for helping to make the department such a pleasant place to
work. Finally I wish to thank Paul Vitanyi and Lambert Meertens for providing me
with the chance to work at the Centrum voor Wiskunde en Informatica in Amsterdam
for the year between my time in Arizona and my move to Oregon, and for
permitting me to finish work on the book while there.

\input{language/getit}
