\secrel{Processing a Class Definition}

Once you have created a class description, the next step is to add the class
to the collection ofstandard classes provided by the Little Smalltalk system.
Suppose, for example, you have defined a class Test in the file test.st; to
add this class definition to a running Little Smalltalk execution you would
type the following command in place of a Smalltalk expression:
\begin{lstlisting}
    )i test.st
\end{lstlisting}

The i can be thought of as mnemonic for "include." At this point the
class description will be parsed, and, if there are syntax errors, a number
of descriptive error messages will be produced. Ifthere are no syntax errors,
the class definition will be added to the collection of classes known to the
Little Smalltalk system. Whether there are syntax errors or not, the cursor
should advance by one tab stop when the system is ready to accept the
next command.

If there are syntax errors, the class description can be edited without
leaving the Little Smalltalk system. To do this, type the following command:
\begin{lstlisting}
    )e test.st
\end{lstlisting}

The )e command will place the user in an editor mode to view the
designated file\note{On UNIX Systems the editor selected can be changed by modifying the EDITOR
environment variable.}. When the user exits the editor, another attempt will be
made to include the file, as if the )i command were typed.

Once a class definition has been successfully included, the defined class
can be used the same as any other class in the Little SmalltaIk system. For
example, a new instance can be created using the command new.
\begin{lstlisting}
    i~Te5t new
\end{lstlisting}

There are other system commands, similar to )i and k, described in
Appendix 1. These commands are provided with an alternative syntax,
rather than as messages passed to smalltalk, because they are used during
the bootstrapping process before any objects have been defined. The bootstrapping of the Little Smalltalk system will be discussed in the second
section of this book.
